---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "README-")
```

# wastdr makes WA Strandings Data accessible in R

The [WA Strandings Database WAStD](https://strandings.dpaw.wa.gov.au/) lives at
[github.com/parksandwildlife/wastd](https://github.com/parksandwildlife/wastd)
and provides a RESTful API.

`wastdr` facilitates reading, parsing and using data from the WAStD API by providing
helpers to access the API as well as examples on how to flatten the GeoJSON from
the API into a two-dimensional tibble.

## Installation

Install `wastdr` from GitHub:

```{r, eval=FALSE}
require(devtools) || install.packages("devtools")
devtools::install_github("parksandwildlife/wastdr")
```


## Setup

Find your valid WAStD API Token at [WAStD](https://strandings.dpaw.wa.gov.au/) 
under "My Profile" and run:

```{r}
wastdr::wastdr_setup(api_token = "c12345asdfqwer")
```

Review the settings with:

```{r}
library(wastdr)
wastdr_settings()
```

## Get WAStD

```{r, eval=FALSE}
track_records <- get_wastd('turtle-nest-encounters')
listviewer::jsonedit(track_records)
```

## Learn more

See the vignette for in-depth examples of transforming and analysing data.

```{r, eval=FALSE}
vignette("getting-wastd")
```
