---
title: "Data analysis - Producing insight from WAStD"
output: 
  rmarkdown::html_vignette:
    self_contained: false
vignette: >
    %\VignetteIndexEntry{analysis}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval=T)
library(wastdr)
library(dplyr)
library(tidyr)
library(skimr)
library(leaflet)
library(lubridate)
library(listviewer)
library(ggplot2)
library(reactable)
wastdr::wastdr_setup()
# wastdr::wastdr_settings()

# A shortcut for reactable
rt <- . %>% reactable::reactable(filterable = TRUE,
                                 sortable = TRUE,
                                 searchable = TRUE,
                                 defaultColDef = colDef(html = TRUE))
```
# Data

## Load data
Download all the data live. This will take a while, but then allow to produce
several reports (e.g. for different turtle programs) at once, or serve as data
for an RShiny app like the turtle data preview.

```{r load_data, eval=F, echo=T}
data_wastd <- download_wastd_turtledata()
```

For a quick example, rendering this vignette, or skipping WAStD authentication,
use the packaged data.

```{r load_data_pkg, echo=T}
data("wastd_data")
```

## Filter data
To filter records to one area, we can either filter by area or site ID , 
or simply filter by a bounding box. Additionally, we'll filter by date.

Site names and IDs are only correct at the time of writing. 
As they could change, please double-check the correct spelling of your own 
place names and their IDs in [WAStD's Areas of type "Locality"](https://wastd.dbca.wa.gov.au/admin/observations/area/?area_type__exact=Locality).

```{r filter_data}
seasons <- wastd_data$tracks %>% 
  dplyr::group_by(season) %>% 
  dplyr::tally() %>% 
  dplyr::arrange(-n) %>% 
  magrittr::extract2("season")

tracks <- wastd_data$tracks %>% filter_wastd_season(seasons[[1]])
```

# Turtle Tracks and Nests
This section demonstrates the summary and visualisation utilities for tracks.

## Maps
Each track on the map has a pop-up with a link to the actual record on WAStD.
Accessing and editing the record is restricted to authorised DBCA staff members.

Map markers can be grouped or not. 

Non-grouped markers show the actual location more accurately, but can overlap to
the point where it is not possible to open the pop-ups. 
Maps with many markers can be slow to render.

Grouped markers in densely populated locations will expand to prevent the 
overlap of markers and pop-ups. Maps with many clustered markers render fast.

Use "Place names" where the aerial imagery is not available at the zoom level 
required for the grouped markers to expand.

```{r tracks_maps}
wastd_data$tracks %>% 
  map_tracks(sites = wastd_data$sites)

# Many tracks display more performant when clustered
wastd_data$tracks %>% map_tracks(sites = wastd_data$sites, cluster=T)
```


## Nesting success - tracks with nest vs the rest
Overview tables are available for the most common summaries at the most common
groupings (season, week, day / area, site).

We show print-optimised `knitr` tables here, but could also use the more refined
`reactable` (filterable, sortable, searchable) for interactive use.

```{r tracks_ns}
wastd_data$tracks %>% nesting_type_by_season_species() %>% rt
wastd_data$tracks %>% nesting_type_by_area_season_species() %>% rt
wastd_data$tracks %>% nesting_type_by_site_season_species() %>% rt
wastd_data$tracks %>% nesting_type_by_season_week_species() %>% rt
wastd_data$tracks %>% nesting_type_by_season_day_species() %>% rt

# Track success by day and species
wastd_data$tracks %>% track_success() %>% rt

# Track success by species
wastd_data$tracks %>% 
  track_success() %>% 
  track_success_by_species() %>% 
  rt

wastd_data$tracks %>% 
  track_success %>% 
  ggplot_track_success_by_date(
    "natator-depressus", placename="Test place", prefix="TEST")

wastd_data$tracks %>% 
  track_success %>% 
  ggplot_track_successrate_by_date(
    "natator-depressus", placename="Test place", prefix="TEST")
```

## Nest excavations: Hatching and emergence success

```{r tracks_hses}
wastd_data$nest_excavations %>% hatching_emergence_success() %>% rt
wastd_data$nest_excavations %>% hatching_emergence_success_area() %>% rt
wastd_data$nest_excavations %>% hatching_emergence_success_site() %>% rt
```

## Tagged nests
```{r tagged_nests}
wastd_data$nest_tags %>% map_nests()
```


## Disturbed nests
```{r dist_nests}
wastd_data$tracks %>% 
  dplyr::filter(disturbance == "present") %>% 
  add_nest_labels() %>% 
  map_tracks(sites = wastd_data$sites)

wastd_data$tracks %>% 
  dplyr::filter(disturbance == "present") %>% 
  add_nest_labels() %>% 
  map_tracks(sites = wastd_data$sites, cluster=T)
```


# Disturbances and predation
```{r}
wastd_data$nest_dist %>% disturbance_by_season() %>% rt
wastd_data$nest_dist %>% map_dist()
wastd_data$nest_dist %>% map_dist(cluster=T)
```

# Animals

## Live sightings: rescues, in water, tagging

```{r}
wastd_data$animals %>% filter_alive() %>% map_mwi(sites = wastd_data$sites)
```

## Dead sightings: strandings, mortalities

```{r}
wastd_data$animals %>% filter_dead() %>% map_mwi(sites = wastd_data$sites)
```

# Surveys

```{r surveys, fig.width=7}
wastd_data$surveys %>% surveys_per_site_name_and_date() %>% rt
wastd_data$surveys %>% survey_hours_per_site_name_and_date() %>% rt
wastd_data$surveys %>% survey_hours_per_person() %>% rt
wastd_data$surveys %>% list_survey_count()
wastd_data$surveys %>% plot_survey_count()
wastd_data$surveys %>% list_survey_effort()
wastd_data$surveys %>% plot_survey_effort()
wastd_data$surveys %>%
  survey_hours_heatmap(placename = "All the places", prefix = "TEST")
wastd_data$surveys %>%
  survey_count_heatmap(placename = "All the places", prefix = "TEST")
wastd_data$surveys %>% survey_season_stats() %>% rt
wastd_data$surveys %>% survey_season_site_stats() %>% rt
wastd_data$surveys %>% survey_show_detail() %>% rt
```

# QA

QA products are part of the ETL pipeline in R package
[etlTurtleNesting](https://github.com/dbca-wa/etlTurtleNesting).

Currently, there are QA reports for user mapping as well as sites and surveys.

