---
title: "Data analysis - Producing insight from WAStD"
author: "Florian Mayer, North West Shelf Flatbacks WA"
date: "`r Sys.time()`"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Data analysis}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
library(wastdr)
library(dplyr)
library(tidyr)
library(magrittr)
library(skimr)
library(leaflet)
library(lubridate)
library(listviewer)
library(DT)
library(ggplot2)
wastdr::wastdr_setup()
# wastdr::wastdr_settings()
```
# Data
This section illustrates how to access data, and defines helper methods to 
filter, map, summarise, and plot data.

Once refined, these helper functions will become core `wastdr` functionality.

## Caveat
* The data as shown is still pre-QA until this message disappears.
* Usernames are still hilariously wrong after matching the usernames 
  (as submitted) to the actual usernames in WAStD. See the field `commments` for
  QA messages.

## Load data
```{r load_data, echo=T}
if (file.exists("tracks.Rda")){
    load("tracks.Rda")
} else {
    track_records <- wastdr::wastd_GET("turtle-nest-encounters")
    tracks_all <- parse_turtle_nest_encounters(track_records)
    surveys <- wastd_GET("surveys") %>% parse_surveys()
    save(tracks_all, track_records, surveys, file = "tracks.Rda")
}
```

## Filter data
To filter records to one area, we can either filter by area or site ID 
(once enabled), or simply filter by a bounding box. 
Additionaly, we'll filter by date.

Site names and IDs are only correct at the time of writing. As they could change, please
double-check the correct spelling of your own place names and their IDs in WAStD.

```{r filter_data}
filter_2017 <- . %>% dplyr::filter(date > dmy("01/10/2017"), date < dmy("31/03/2018"))
filter_broome <- . %>% dplyr::filter(area_name=="Cable Beach Broome")
filter_broome_sites <- . %>% dplyr::filter(site_id %in% c(22, 23, 24))
filter_cbb1 <- . %>% dplyr::filter(site_name=="Cable Beach Broome Sector 1")
filter_cbb2 <- . %>% dplyr::filter(site_name=="Cable Beach Broome Sector 2")
filter_cbb3 <- . %>% dplyr::filter(site_name=="Cable Beach Broome Sector 3")
filter_eighty_mile_beach <- . %>% dplyr::filter(area_name=="Eighty Mile Beach Caravan Park")
filter_anna_plains <- . %>% dplyr::filter(area_name=="Anna Plains")
filter_port_hedland_sites <- . %>% dplyr::filter(site_id %in% c(35, 45))
filter_port_hedland_cemetery <- . %>% dplyr::filter(site_name=="Port Hedland Cemetery Beach")
filter_port_hedland_prettypool <- . %>% dplyr::filter(site_name=="Port Hedland Pretty Pool Beach")
filter_west_pilbara <- . %>% dplyr::filter(area_name=="West Pilbara Turtle Program beaches Wickam")
filter_delambre <- . %>% dplyr::filter(area_name=="Delambre Island")
filter_rosemary <- . %>% dplyr::filter(area_name=="Rosemary Island")
filter_thevenard <- . %>% dplyr::filter(area_name=="Thevenard Island")

tracks <- tracks_all %>% filter_2017
```


## Map data

```{r, eval=T}
tracks_map <- function(track_data) {
    tracks <- track_data %>% add_nest_labels
    
    l <- leaflet::leaflet(width=1000, height=800) %>% 
        addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
        addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
        clearBounds()

    tracks.df <-  tracks %>% split(tracks$species)
    
    names(tracks.df) %>%
        purrr::walk( function(df) {
            l <<- l %>%
                addAwesomeMarkers(
                    data = tracks.df[[df]],
                    lng = ~longitude, lat=~latitude,
                    icon = leaflet::makeAwesomeIcon(
                        text = ~nest_type_text,
                        markerColor = ~species_colours),
                    label=~paste(date, nest_age, species, nest_type, name),
                    popup=~paste(date, nest_age, species, nest_type, name),
                    group = df
                )
        })
    
    l %>%
        addLayersControl(
            baseGroups = c("Aerial", "Place names"),
            overlayGroups = names(tracks.df),
            options = layersControlOptions(collapsed = FALSE)
        )
}
```


## Summarise and plot data
Data (all tracks or filtered subsets) are filtered to only fresh observations,
then grouped and tallied by date, species and type.

Daily summaries are shown in wide form as tables, and (using long form) as 
timeseries plots.

```{r data_wp_daily}
species_by_type <- . %>% 
  filter(nest_age=="fresh") %>%
  group_by(species, nest_type) %>% 
  tally() %>%
  ungroup() %>% 
  tidyr::spread(nest_type, n, fill=0)

daily_species_by_type <- . %>% 
    filter(nest_age=="fresh") %>%
    group_by(date, species, nest_type) %>% 
    tally() %>%
    ungroup()

daily_summary <- . %>% 
    daily_species_by_type %>% 
    tidyr::spread(nest_type, n, fill=0) %>%
    DT::datatable(.)

tracks_ts <- . %>% 
    daily_species_by_type %>% 
    {ggplot2::ggplot(data=., aes(x = date, y = n, colour = nest_type)) + 
            ggplot2::geom_point() + 
            ggplot2::geom_smooth(method = "auto") +
            # ggplot2::geom_line() +
            ggplot2::scale_x_date(breaks = scales::pretty_breaks(),
                                  labels = scales::date_format("%d %b %Y")) +
            ggplot2::scale_y_continuous(limits = c(0, NA)) +
            ggplot2::xlab("Date") +
            ggplot2::ylab("Number counted per day") +
            ggplot2::ggtitle("Nesting activity") +
            ggplot2::theme_light()}
```

# Nesting activity
This chapter uses the data and helpers from the above section and provides some
insight into the different regions.

This section is by no means complete and can be extended as appropriate.

## Overall
```{r, fig.width=7, fig.height=5}
tracks %>% tracks_map
```

## Broome Cable Beach 
```{r, fig.width=9, fig.height=5}
tracks_cbb <- tracks %>% filter_broome
tracks_cbb %>% tracks_map
tracks_cbb %>% DT::datatable(.)
tracks_cbb %>% daily_summary
tracks_cbb %>% tracks_ts

# named_nests_cbb <- tracks_cbb %>% filter(!(is.na(name)))
# named_nests_cbb %>% tracks_map
# named_nests_cbb %>% DT::datatable(.)

surveys_cbb <- surveys %>% filter_broome_sites
place <- "Cable Beach Broome"
surveys_cbb %>% plot_survey_count(place)
surveys_cbb %>% list_survey_count(place)
surveys_cbb %>% plot_survey_effort(place)
surveys_cbb %>% list_survey_effort(place)
```

## Anna Plains
```{r, fig.width=7, fig.height=5}
tracks_ap <- tracks %>% filter_anna_plains
tracks_ap %>% tracks_map
tracks_ap %>% DT::datatable(.)
tracks_ap %>% daily_summary
tracks_ap %>% tracks_ts
```

## Eighty Mile Beach
```{r, fig.width=7, fig.height=5}
tracks_emb <- tracks %>% filter_eighty_mile_beach
tracks_emb %>% tracks_map
tracks_emb %>% DT::datatable(.)
tracks_emb %>% daily_summary
tracks_emb %>% tracks_ts
```

## Port Hedland

```{r, fig.width=9, fig.height=5}
tracks_pth_cem <- tracks %>% filter_port_hedland_cemetery
tracks_pth_ppo <- tracks %>% filter_port_hedland_prettypool

tracks_pth_cem %>% tracks_map
tracks_pth_cem %>% DT::datatable(.)
tracks_pth_cem %>% daily_summary
tracks_pth_cem %>% tracks_ts %T>% ggsave(filename="~/pth_daily_tracks_cem.png", device="png", width=9, height=5)

tracks_pth_ppo %>% tracks_map
tracks_pth_ppo %>% DT::datatable(.)
tracks_pth_ppo %>% daily_summary
tracks_pth_ppo %>% tracks_ts %T>% ggsave(filename="~/pth_daily_tracks_ppo.png", device="png", width=9, height=5)

surveys_pth <- surveys %>% filter_port_hedland_sites
place <- "Port Hedland"
surveys_pth %>% plot_survey_count(place)
surveys_pth %>% list_survey_count(place)
surveys_pth %>% plot_survey_effort(place)
surveys_pth %>% list_survey_effort(place)

# named_nests_pth <- tracks_pth %>% filter(!(is.na(name)))
# named_nests_pth %>% tracks_map
# named_nests_pth %>% DT::datatable(.)
```

## West Pilbara Turtle Program

```{r, fig.width=7, fig.height=5}
tracks_wp <- tracks %>% filter_west_pilbara
tracks_wp %>% tracks_map
tracks_wp %>% DT::datatable(.)
tracks_wp %>% daily_summary
tracks_wp %>% tracks_ts
```

## Delambre Island

```{r, fig.width=7, fig.height=5}
tracks_de <- tracks %>% filter_delambre
tracks_de %>% tracks_map
tracks_de %>% DT::datatable(.)
tracks_de %>% daily_summary
tracks_de %>% tracks_ts
```

## Rosemary Island
Data pending upload from tablets.

```{r, fig.width=7, fig.height=5, eval=F}
tracks_ri <- tracks %>% filter_rosemary
tracks_ri %>% tracks_map
tracks_ri %>% DT::datatable(.)
tracks_ri %>% daily_summary
tracks_ri %>% tracks_ts
```

## Thevenard Island

```{r, fig.width=7, fig.height=5}
tracks_thv <- tracks %>% filter_thevenard
tracks_thv %>% tracks_map
tracks_thv %>% DT::datatable(.)
tracks_thv %>% daily_summary
tracks_thv %>% tracks_ts
```
