---
title: "Getting WAStD"
author: "Florian Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting WAStD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup

Find your valid WAStD API Token at [WAStD](https://strandings.dpaw.wa.gov.au/) 
under "My Profile" and run:

```{r setup}
require(dplyr)
require(lubridate)
require(stringr)
require(tidyverse)
require(magrittr)
require(listviewer)
library(wastdr)
wastdr::wastdr_setup(api_token = "c12345asdfqwer")
```

Review the `wastdr` settings with:

```{r review_settings}
wastdr::wastdr_settings()
```

## Get WAStD

With your valid API key and access to the WAStD API, you can retrieve data with
`get_wastd`.

```{r call_api, eval=FALSE}
tag_json <- get_wastd("animal-encounters", 
                      query=list(taxon="Cheloniidae", 
                                 limit=300, 
                                 format="json"))
listviewer::jsonedit(head(tag_json$content))
tags <- tag_json %>% parse_animal_encounters
# devtools::use_data(tags, overwrite=T)

track_json <- get_wastd("turtle-nest-encounters", 
                        query=list(taxon="Cheloniidae", 
                                   limit=300, 
                                   format="json"))
tracks <- nest_json %>% parse_turtle_nest_encounters
listviewer::jsonedit(head(track_json$content))
# devtools::use_data(tracks, overwrite=T)

nest_json <- get_wastd("turtle-nest-encounters", 
                       query=list(taxon="Cheloniidae", 
                                  limit=300, 
                                  format="json", 
                                  nest_type="hatched-nest"))
nests <- nest_json %>% parse_turtle_nest_encounters
listviewer::jsonedit(head(nest_json$content))
# devtools::use_data(nests, overwrite=T)
```

For this vignette, we'll use the packaged example data.

```{r}
data("tags")
data("tracks")
data("nests")
```


## Formatting the response

TODO split out functions, write tests

```{r, eval=F}
#' Filter a data.frame to records collected at THV after 19 Dec 2016
#' TODO replace with filtering for THV 2016/2017 field trips at API
#' with query parameter `expedition=1`.
thv_filter <- . %>% filter(
    latitude < -21.43,
    latitude > -21.48,
    longitude > 114.96,
    longitude < 115.05,
    date > dmy("17/11/2016"))

tracks <- tracks %>% thv_filter

nests <- nest_json %>% parse_turtle_nest_encounters %>% thv_filter


```


## Visualise

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

