---
title: "Getting WAStD"
author: "Florian Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting WAStD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Get WAStD

After setting up `wastdr`'s authentication in vignette "Setup", 
we can now retrieve data using `wastdr::wastd_GET()` to send `GET` 
requests to the `WAStD` API.

This vignette shows examples of retrieving and parsing data from all 
existing WAStD API endpoints. The intent of this vignette is to serve as a 
starting point

See vignette "get-tsc" for examples of the Threatened Species and Communities side. 

```{r}
library(wastdr)
library(dplyr)
library(magrittr)
library(leaflet)
library(listviewer)
library(DT)
library(ggplot2)
```



# Encounters
```{r}
# Encounters from 2016 onwards (this includes disturbances, animal and turtle 
# nest encounters but not their specific details)
# https://tsc.dbca.wa.gov.au/api/1/animal-encounters/?when__year__gte=2016

```

## Filtering the results
See examples of filter parameters and possibl efilter values on:

* [https://tsc.dbca.wa.gov.au/api/1/animal-encounters/](https://tsc.dbca.wa.gov.au/api/1/animal-encounters/)
* []()
* []()

This section will be expanded to show how to filter the data, e.g.

* All data from an area (marine protected area, location, study site)
* All data collected or recorded by a certain person
* All data with specific properties, e.g. only certain stranding types, taxonomic groups etc.

## AnimalEncounters
```{r}
# AnimalEncounters of Turtles from 2016 onwards
# https://tsc.dbca.wa.gov.au/api/1/animal-encounters/?taxon=Cheloniidae&when__year__gte=2016
```

## TurtleNestEncounters

```{r}
# https://tsc.dbca.wa.gov.au/api/1/turtle-nest-encounters/?taxon=Cheloniidae&when__year__gte=2016
```


## LineTransectEncounters

## LoggerEncounters



