---
title: "Getting WAStD"
author: "Florian Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting WAStD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup

Find your valid WAStD API Token at [WAStD](https://strandings.dpaw.wa.gov.au/) 
under "My Profile" and run:

```{r setup}
require(dplyr)
require(lubridate)
require(stringr)
require(tidyverse)
require(magrittr)
require(listviewer)
library(wastdr)
wastdr::wastdr_setup(api_token = "c12345asdfqwer")
```

Review the `wastdr` settings with:

```{r review_settings}
wastdr::wastdr_settings()
```

## Get WAStD

With your valid API key and access to the WAStD API, you can retrieve data with
`get_wastd`.

```{r call_api, eval=FALSE}
tag_json <- get_wastd("animal-encounters", 
                             query=list(taxon="Cheloniidae", 
                                        limit=300, 
                                        format="json"))
track_json <- get_wastd("turtle-nest-encounters", 
                             query=list(taxon="Cheloniidae", 
                                        limit=300, 
                                        format="json"))
nest_json <- get_wastd("turtle-nest-encounters", 
                             query=list(taxon="Cheloniidae", 
                                        limit=300, 
                                        format="json", 
                                        nest_type="hatched-nest"))
```

For this vignette, we'll use the packaged data.

```{r}
tag_json <- data("tag_json")
track_json <- data("track_json")
nest_json <- data("nest_json")
```

```{r}
# listviewer::jsonedit(head(nest_json$content))
```

## Formatting the response

TODO split out functions, write tests

```{r}
#' Convert between url-safe dash-separated-names and Human Readable Title Case
humanize <- . %>% stringr::str_to_title() %>% stringr::str_replace("-", " ")
urlize <- . %>% stringr::str_to_lower() %>% stringr::str_replace(" ", "-")

#' O Lord of Darkness, accept this our functional sacrifice of three days and three nights
#' From a list of unnamed lists, or the data.frame equivalent, extract a field or -1
extract_possibly <- purrr::possibly(magrittr::extract, otherwise = -1)
get_f <- function(lol, field) lol %>% unlist %>% extract_possibly(field) %>% unlist %>% as.numeric

#' Filter a data.frame to records collected at THV after 19 Dec 2016
#' TODO replace with filtering for THV 2016/2017 field trips at API
thv_filter <- . %>% filter(
    latitude < -21.43,
    latitude > -21.48,
    longitude > 114.96,
    longitude < 115.05,
    date > dmy("17/11/2016"))

# nests <- nest_json$content %>% {
#   tibble::tibble(
#     datetime = map_chr(., c("properties", "when")) %>% utc_as_gmt08,
#     longitude = map_dbl(., c("properties", "longitude")),
#     latitude = map_dbl(., c("properties", "latitude")),
#     date = map_chr(., c("properties", "when")) %>% as_turtle_date,
#     species = map_chr(., c("properties", "species")),
#     obs = map(., c("properties", "observation_set")),
#     hatching_success = obs %>% map(get_f, "hatching_success") %>% as.numeric,
#     emergence_success = obs %>% map(get_f, "emergence_success") %>% as.numeric,
#     clutch_size = obs %>% map(get_f, "egg_count_calculated") %>% as.numeric
#     )
# } %>% thv_filter
```


## Visualise

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

