---
title: "Setup"
author: "Florian Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = T,
  eval = T,
  comment = "#>"
)
```
# Authentication

This vignette explains how to authenticate to the WAStD API.

```{r, message=FALSE}
require(wastdr)
```
There are two audiences with two different authentication options.

The first audience, DBCA staff, has access to 
[WAStD](https://strandings.dpaw.wa.gov.au).
On their profile page they will find their own API token.

The second audience, external collaborators, will be given a username and password.
Using these credentials, they can access the 
[WAStD API](https://strandings.dpaw.wa.gov.au/api/1/) 
both through their browser and from other software like `wastdr`.

`wastdr::get_wastd()` accesses the `WAStD` API using:

* the API URL, and 
* an access token **or** a username / password.

There are three ways to supply the required settings to `wastdr::get_wastd()`.


## Best practice: `.Renviron`

For a permanent setup, the environment variables can be set in the user's
`.Renviron`.

DBCA staff add to their `.Renviron`:

```{r, eval=F}
WASTDR_API_URL="https://tsc.dbca.wa.gov.au/api/1/"
WASTDR_API_UN="XXX"
WASTDR_API_PW="XXX"
```

Note that the API token contains the word `Token` and a whitespace in addition
to the actual API token.

External collaborators add to their `.Renviron`:

```{r, eval=F}
WASTDR_API_URL="https://tsc.dbca.wa.gov.au/api/1/"
WASTDR_API_UN="XXX"
WASTDR_API_PW="XXX"
```

In the long term, this last method is the most elegant and persistent way of
configuring `wastdr`.

## Per session

If `api_*` settings are not provided, `wastdr::get_wastd` defaults to 
specifically named environment variables.

For convenience, `wastdr_setup` sets these environment variables, while 
`wastdr_settings` retrieves the currently set values.

If these variables haven't been provided through e.g. `.Renviron`, you may set 
them on a per session basis:

```{r, eval=F}
wastdr::wastdr_setup(api_token = "Token XXX")
```

Alternatively you may use your username and password.

```{r, eval=F}
wastdr::wastdr_setup(api_un = "XXX", api_pw = "XXX")
```


Review the settings with:

```{r}
wastdr::wastdr_settings()
```

`wastdr_setup` defaults `api_url` to the currently only existing API url.


## Per function call
Crendentials can be supplied directly to functions:

```{r eval=F}
wastdr::wastd_GET('encounters', 
                  api_token = "Token XXX")

wastdr::wastd_GET('encounters',
                  api_un = "XXX",
                  api_pw = "XXX")
```

This method overrides any other settings, but leads to verbose code with confidential
data (token or password) in plain text.

