---
title: "Setup"
author: "Florian Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```
# Authentication

This vignette explains how to authenticate to the WAStD API.

```{r, message=FALSE}
require(wastdr)
```
There are two audiences with two different authentication options.

The first audience, DBCA staff, has access to 
[WAStD](https://strandings.dpaw.wa.gov.au).
On their profile page they will find their own API token.

The second audience, external collaborators, will be given a username and password.
Using these credentials, they can access the 
[WAStD API](https://strandings.dpaw.wa.gov.au/api/1/) 
both through their browser and from other software like `wastdr`.

# Configure credentials
`wastdr::get_wastd()` accesses the `WAStD` API using:

* the API URL, and 
* an access token **or** a username / password.

There are three ways to supply the required settings to `wastdr::get_wastd()`.

## Per function call
Firstly, they can be supplied directly to `wastdr::get_wastd()`:

```{r eval=F}
wastdr::get_wastd('encounters', 
                  api_token = "asdfqwer1234")

wastdr::get_wastd('encounters',
                  api_un = "my_usernname",
                  api_pw = "my_password")
```

This method overrides any other settings, but leads to verbose code with confidential
data (token or password) in plain text.

## Per session

If `api_*` settings are not provided, `wastdr::get_wastd` defaults to 
specifically named environment variables.

For convenience, `wastdr_setup` sets these environment variables, while 
`wastdr_settings` retrieves the currently set values.

If you are an employee of the Western Australian Department of Biodiversity,
Conservation and Attractions (DBCA), formerly the Department of Parks and Wildlife,
then you can find your valid WAStD API Token at [WAStD](https://strandings.dpaw.wa.gov.au/) 
under "My Profile" and run:

```{r}
wastdr::wastdr_setup(api_token = "c12345asdfqwer")
```

If you are an external collaborator, you will have given a username and password
to access the WAStD API.

```{r}
wastdr::wastdr_setup(api_un = "my_username", api_pw = "my_password")
```


Review the settings with:

```{r}
wastdr::wastdr_settings()
```

`wastdr_setup` defaults `api_url` to the currently only existing API url.

This method reduces verbosity, but still exposes sensitive data in the `wastdr_setup` call.
The `wastdr::wastdr_settings()` expire with the R session.

## Per machine

For a more permanent setup, the environment variables can be set in the user's
`.Rprofile`.

DBCA staff add to their `.Rprofile`:

```{r}
Sys.setenv(WASTDR_API_TOKEN="Token c1234")
```

Note that the API token contains the word `Token` and a whitespace in addition
to the actual API token.

External collaborators add to their `.Rprofile`:

```{r}
Sys.setenv(WASTDR_API_UN="my_username")
Sys.setenv(WASTDR_API_PW="my_password")
```

In the long term, this last method is the most elegant and persistent way of
configuring `wastdr`.
