---
title: "Turtle Tagging ETL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Turtle Tagging ETL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wastdr)
```

This workflow copies the contents of the turtle tagging database WAMTRAM2 into
the WA Sea Turtle Database (WAStD).

* Load the turtle tagging db WAMTRAM from its MS SQL Server database into memory 
  and save objects as RData
* Load existing tagging records from TSC: load only a minimal set of 
  source, source ID, QA status to determine later what to create / update / skip:
  * does not exist in TSC: create (POST)
  * exists in TSC with status "new": update (PATCH)
  * exists in TSC with status higher than "new": skip (and message)
* Make (transform) WAMTRAM to TSC data
* Load transformed data into TSC's API (create/update/skip)
* Extract tagging records from TSC again
* Show (map, tabulate, summarise) tagging records from TSC
* Compare tagging data from TSC to tagging data from WAMTRAM - 
  e.g. tally across factor levels (species, health, tag status). 
  At least compare: same number of sightings of same tag?


