---
title: "WAStD UserMapping QA"
output: html_document
---
<style type="text/css">
.main-container {
  max-width: 100%;
  margin-left: auto;
  margin-right: auto;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pointblank)
library(reactable)
library(drake)
# drake::loadd("user_mapping")
```

# WAStD User Mapping
## Step 0: Add new data collectors to WAStD
Coordinators of data capture programs supply us with a spreadsheet in this exact format:

* Column "name": The full name, including given and surname, e.g. "Florian Mayer"
  * Use correct capitalisation. Do not write "florian mayer" or "FLORIAN MAYER".
    The case does not matter for the user matching, but it matters for displaying
    in reports.
* Column "email": A valid email, e.g. "Florian.Mayer@dbca.wa.gov.au". 
  * Omit any HTML markup, e.g. "mailto:", or anything involving `<`pointy brackets`>`.
* Column "phone": A valid phone number with the Australian prefix `+61`.
  * Valid example: "+61414123456".
  * Whitespace will be discarded, so "+61 414 123 456" works too.
  * Format the column as "text". Excel has a nasty habit of auto-formatting
    columns and dropping the leading "+". This is very tedious to fix.
    E.g. "61414123456" is invalid.

We add columns:

* "username", generated from the snake-cased "name", e.g. "florian_mayer".
* "role", which we set to a predefined text such as "Cable Beach Broome Volunteer 2020-21".
  As user registrations come in batches from distinct locations, we get a cleaner
  result if we set the role centrally following a naming convention.

## Step 1: wastdr takes a wild guess
During the data import from ODK Central to WAStD, wastdr maps each distinct 
ODK Collect "username" as written by data collectors to actual WAStD user profiles.

The matching is done by 
[`fuzzyjoin::stringdist_left_join`](http://varianceexplained.org/fuzzyjoin/reference/stringdist_join.html) using the Jaro-Winker distance between the ODKC username and the combined WAStD `name` and `aliases`. 

Details about the distance measures are [here](http://finzi.psych.upenn.edu/library/stringdist/html/stringdist-metrics.html).

The Jaro-Winker distance was chosen as it returns the highest number of correct matches.

## Step 2: Manually update WAStD user aliases to improve mapping
In the automated validation output we show all user mappings with a dissimilarity
above a threshold. This should separate mismatches from matches.

The validation table shows columns:

* `odkc_username` - the ODK Collect username
* `wastd_matched` - a link to the WAStD User profile that we matched
* `search_wastd` - individual links to search WAStD Users (names, usernames, 
  aliases, role, phone, email) for each part of the ODK Collect username.
  This is helpful where a list of names were supplied, e.g. "Flo and Janet".
  
For each mismatch:
Find the correct WAStD user, and add the `odkc_username` to the WAStD User's field
"aliases", then "save" or "save and continue editing" the WAStD user profile.
It is **absolutely essential** to separate the aliases with a **comma**.
e.g. Florian Mayer's aliases can be "Flo Mayer, FlorianM, FloJo", which then will
precisely map to "Flo Mayer", "FlorianM", and "FloJo".

Note: updating WAStD will not refresh this report automatically - we have to re-run
the data import.

When we re-run the user matching with fresh data, this user should match up. 
Some will won't, which we'll handle in the next step.

You can download a spreadsheet of mismatches from the CSV button:

```{r message = FALSE, echo=FALSE, validate = TRUE}
a <- user_mapping %>%
    wastdr::annotate_user_mapping() %>% 
    pointblank::create_agent() %>%
    pointblank::col_vals_lt("dist", 0.10) %>%
    interrogate()
a
```


Alternatively, you can work through the issues right here:

```{r echo=FALSE, validate=TRUE}
a %>%
  pointblank::get_data_extracts() %>%
  magrittr::extract2(1) %>%
  gt::gt() %>%
  gt::fmt_markdown(columns = TRUE) %>% 
  gt::cols_label(
    odkc_username = gt::html("<h5>They wrote</h5>\n<small>ODK Collect username</small>"),
    wastd_matched = gt::html("<h5>We matched</h5>\n<small>WAStD User</small>"),
    search_wastd = gt::html("<h5>Search</h5>\n<small>likely candidates</small>"),
    dist = gt::html("<h5>Dissimilarity</h5>\n<small>smaller = better</small>")
  ) %>% 
  gt::tab_spanner(
    label = "WASTD User match",
    columns = vars(odkc_username, wastd_matched, search_wastd, dist)
  ) %>% 
  gt::tab_spanner(
    label = "WASTD User profile",
    columns = vars(role, pk, username, name, nickname, aliases, email)
  )
```

